### 2014-10-10 LaTeX/Abstract.tex
More changes.

% Version submitted to MPSA:
Within the conflict literature there has been much disagreement about the relationship between civil wars, natural resources, and state economic performance. We find that this disagreement results from not accounting for the spatial disaggregation of conflict events within a country. Our hypothesis is that the economic impact of civil conflict is contingent on the conflict's location relative to major economic and labor resources within a state. We use subnational data on the spatial distribution of conflict, resources, and infrastructure to test the long-term impact of domestic conflict on state economic performance. To estimate the spatial distribution of conflict we use data from the PRIO Armed Conflict Location and Event Data. We combine the conflict location data with geospatial data on economic centers, natural resource locations, and infrastructure grids to generate spatial variables that approximate how far each conflict is from centers of interest. We then estimate the effect of the spatial variables on economic performance. By doing so, we are able to resolve tensions in the literature regarding the relationship between economic performance and civil wars.

>>>>

% Version submitted to MPSA:
Within the conflict literature there has been much disagreement about the relationship between civil wars and state economic performance. While civil war is often associated with poor economic performance, some states have managed to weather domestic armed conflict during periods of robust growth. We find that this disagreement results from not accounting for the spatial distribution of conflict within a country. Literature in economics has recently stressed the importance of major cities to development and economic growth. We hypothesize that the economic impact of civil conflict is contingent on the conflict's location relative to major urban centers within a state. We use subnational data on the location of conflict relative to urban areas to test the impact of domestic conflict on annual GDP growth. In doing so, we bridge the economic development literature on the importance of cities with extant literature on the effect of armed conflict to provide a novel explanation for the paradox of high macroeconomic growth in conflict ridden countries. By doing so, we are able to resolve tensions in the literature regarding the relationship between economic performance and civil wars. We find that the proximity of civil war to major urban centers is negatively correlated with economic growth and use simulations to show that the estimated marginal effect of civil war proximity can mean the difference between robust growth and stagnation.

### 2014-10-10 LaTeX/RhodiumEmp.tex
More changes.

We collect information on the location of conflicts from the PRIO Conflict Site Dataset \citep{hallberg:2012}. This dataset contains information on the geographical centers of armed conflict events and the area covered by each conflict at a yearly level from 1989 to 2008. In recent years, other geo-referenced datasets have been developed that go beyond just providing information on conflict centroids and area to providing an actual spatial grid structure through which to conduct analysis (e.g., \citealt{tollefsen:etal:2012}). For the purpose of this analysis, however, we deliberately avoid using these datasets since our primary interest is in examining variation in changes to GDP growth due to the proximity of conflict to urban centers.

% Note to Ben: Does that sound convincing? Probably not. The point that I'm trying to get across is that using the gridded data is not at the appropriate level of analysis. We need to say something here about this so if you can think of something better that'd be great.

To construct a time-series cross-sectional database of urban centers, we turn to yearly editions of \emph{The World Almanac} from 1989 to 2008. More refined subnational data listing urban centers by their contribution to economic output would be the most direct way to test our hypothesis, however, cross-national data such as this over the time period of our analysis, especially for developing countries, is simply not available. The World Alamanac lists the major cities in a country by population, thus providing us with a second best approximation of the urban centers that are most relevant for a particular national economy. Typically, the Almanacs list at least three major cities, including the capital, for each country and year from 1989 to 2008. We use the top five major cities listed for each country and then determine their geographical centers.

@@

% Note to Ben: Does empirical model sound weird?

@@

From the PRIO Armed Conflict dataset we bring in additional information about conflict intensity, duration, the area covered by the conflict, and the number of conflicts \citep{themner:wallensteen:2014}. Conflict intensity is a binary that distinguishes between minor armed conflicts and wars, and we set this equal to one if any of the conflicts in a country year were classified as a war and zero if not. In aggregating conflict duration to the country year level, we bring in the duration (measured in years) of the longest lasting conflict by a certain country-year. Conflict area measures the conflict zone in square kilometers. To include this in our country-year model, we sum up the area covered by each conflict being fought and divide by the total land area of the country. From this we then create a binary measure that is one if greater than 50\% of a country's total land area is immersed in conflict and zero otherwise. Finally, the number of conflicts is simply a count of the number of ongoing internal armed conflicts in a country-year according to the PRIO dataset.

@@

We also include a number of variables that help to capture structural and institutional aspects of the country itself. First, we include a binary for whether the country is classified as upper income by the World Bank and for the level of inflation in the country. Second, we also control for the level of democracy in the country using the polity index \citep{marshall:etal:2013}. We also control for the proportion of a country's GDP that is made up of natural resource rents. This is a particularly important variable to control for as countries that draw a large share of their GDP from natural resource rents likely rely less on urban centers to serve as drivers of growth. Finally, given that the timeframe of our sample overlaps with the occurence of two major global crises, we add a control for the average GDP growth across all countries in the world.

% Note to Ben: Anything else to add about why including polity would be important in the model?

@@

\begin{align*}
	\% \Delta GDP_{i,t} &= \beta_{1}(Ln(Min. \; Conflict \; Dist.)_{i,t-1}) \\
	& \;+\; \beta_{2}(Conflict \; Intensity_{i,t-1}) \;+\; \beta_{3}(Conflict \; Duration_{i,t-1}) \\
	& \;+\; \beta_{4}(Conflict \; Area_{i,t-1}) \;+\; \beta_{5}(Upper \; Income_{i,t}) \\	
	& \;+\; \beta_{6}(Inflation_{i,t-1}) \;+\; \beta_{7}(Democracy_{i,t-1}) \\
	& \;+\; \beta_{8}(Resource \; Rents/GDP_{t-1}) \;+\; \beta_{9}(World \; GDP \; Growth_{t}) \\
	& \;+\; \mu_{i,t} \;+\; \epsilon_{i,t}
\end{align*}

@@

To ensure that our parameter estimates are robust to changes in our sample, we run a six-fold cross-validation. This analysis helps us to understand whether some of the subsets in our dataset follow a different pattern than what is in the broader set \citep{beck2008time}. To conduct the cross-validation, we randomly split the 68 country observations in our dataset into six approximately equal subsets. Each subset ends up containing a minimum of approximately 65 cases. We then run each model shown in figure \ref{fig:coefplot} seven times, where in each iteration we left out one subsample. The results of this analysis are depicted in \ref{fig:crossPlot}, and we can clearly see that the parameter estimates for our two spatial proximity variables remain consistent across the exclusion of any of the folds.

>>>>

We collect information on the location of conflicts from the PRIO Conflict Site Dataset \citep{hallberg:2012}. This dataset contains information on the geographical centers of armed conflict events and the area covered by each conflict at a yearly level from 1989 to 2008. In recent years, other geo-referenced datasets have been developed that go beyond just providing information on conflict centroids and area to providing an actual spatial grid structure through which to conduct analysis \citep{tollefsen:etal:2012}. However, these data are not available globally and would therefore unnecessarily limit the external validity of the current analysis. Other machine coded event datasets, while providing detailed coverage of events on a global scale, often approximate event locations based on descriptors in the source material or by the location of the reporting news agency itself. Because the location of conflict is critical for this analysis, the urban bias of disaggregated event data is unacceptable. For the purpose of this analysis, we deliberately avoid using these datasets since our primary interest is in examining variation in changes to GDP growth at a national level.

To construct a time-series cross-sectional database of urban centers, we turn to yearly editions of \citet{worldalmanac:1989}. More refined subnational data listing urban centers by their contribution to economic output would be the most direct way to test our hypothesis, however, cross-national data such as this over the time period of our analysis, especially for developing countries, is simply not available. \emph{The World Alamanac} lists the major cities in a country by population, thus providing us with a second best approximation of the urban centers that are most relevant for a particular national economy. Typically, the \emph{Almanacs} list at least three major cities, including the capital, for each country and year from 1989 to 2008. Because the \emph{Almanacs} are not perfectly consistent from year to year, we opted to code cities with a five year rolling window. That is, cities listed in the 1995 \emph{Almanac} are also included in our dataset for years 1993-1997. In doing so, we hope to minimize the effect of coding inconsistencies on the part of \emph{The World Almanac}.

@@

% Note to Ben: Does empirical model sound weird?
% Note to Shahryar: It sounds a lot better than "theory."

@@

From the PRIO Armed Conflict dataset we bring in additional information about conflict intensity, duration, the area covered by the conflict, and the number of conflicts \citep{themner:wallensteen:2014}. Conflict intensity is a binary that distinguishes between minor armed conflicts (between 25 and 999 annual battle deaths) and wars (at least 1000 annual battle deaths), and we set this equal to one if any of the conflicts in a country year were classified as a war and zero if not. In aggregating conflict duration to the country year level, we bring in the duration (measured in years) of the longest lasting conflict by a certain country-year. Conflict area measures the conflict zone in square kilometers. To include this in our country-year model, we sum up the area covered by each conflict being fought and divide by the total land area of the country. From this we then create a binary measure that is one if greater than 50\% of a country's total land area is immersed in conflict and zero otherwise. Finally, the number of conflicts is simply a count of the number of ongoing internal armed conflicts in a country-year according to the PRIO dataset.

@@

We also include a number of variables that help to capture structural and institutional aspects of the country itself. First, we include a binary for whether the country is classified as upper income by the World Bank and for the level of inflation in the country. Second, we also control for the level of democracy in the country using the polity index \citep{marshall:etal:2013}. We also control for the proportion of a country's GDP that is made up of natural resource rents. This is a particularly important variable to control for as countries that draw a large share of their GDP from natural resource rents likely rely less on urban centers to serve as drivers of growth. Finally, given that the timeframe of our sample overlaps with the occurrence of two major global economic crises, the 1997 Asian financial crisis and the 2008 financial crisis, we add a control for the average GDP growth across all countries in the world.

% Note to Ben: Anything else to add about why including polity would be important in the model?
% Note to Shahryar: Nope.

@@

\begin{align*}
	\% \Delta GDP_{i,t} &= \beta_{1}(Ln(Min. \; Conflict \; Dist.)_{i,t-1}) \\
	& \;+\; \beta_{2}(Conflict \; Intensity_{i,t-1}) \;+\; \beta_{3}(Conflict \; Duration_{i,t-1}) \\
	& \;+\; \beta_{4}(Conflict \; Area_{i,t-1} \;/\; Land \; Area_{i,t-1}) \\
	& \;+\; \beta_{5}(Number \; of \; Conflicts_{i,t-1}) \;+\; \beta_{6}(Upper \; Income_{i,t}) \\	
	& \;+\; \beta_{7}(Inflation_{i,t-1}) \;+\; \beta_{8}(Democracy_{i,t-1}) \\
	& \;+\; \beta_{9}(Resource \; Rents/GDP_{t-1}) \;+\; \beta_{10}(World \; GDP \; Growth_{t}) \\
	& \;+\; \mu_{i,t} \;+\; \epsilon_{i,t}
\end{align*}

@@

To ensure that our parameter estimates are robust to changes in our sample, we run a six-fold cross-validation. This analysis helps us to understand whether some of the subsets in our dataset follow a different pattern than what is in the broader set \citep{beck2008time}. To conduct the cross-validation, we randomly split the 68 country observations in our dataset into seven approximately equal subsets. Each subset ends up containing a minimum of approximately 65 cases. We then run each model shown in figure \ref{fig:coefplot} seven times, where in each iteration we left out one subsample. The results of this analysis are depicted in \ref{fig:crossPlot}, and we can clearly see that the parameter estimates for our two spatial proximity variables remain consistent across the exclusion of any of the folds.

### 2014-10-10 LaTeX/RhodiumIntro.tex
More changes.

What happens when cities themselves are threatened? In this paper, we bridge the economic development literature on the importance of cities with extant literature on the effect of armed conflict to provide a novel explanation for the paradox of high macroeconomic growth in conflict ridden countries. Specifically,  we argue that to explain variation in economic performance between countries in the midst of civil war, it is necessary to account for the spatial location of the conflict relative to major population and economic centers. Much of the extant literature has focused on examining the effect of conflict on economic growth at the national level. The implicit assumption of these approaches is that all civil wars are alike, whether they occur right outside the capital or on the fringes of the country. We argue that the location of conflicts relative to major cities and capitals will determine the effect of civil conflict on economic performance, and that conflicts only significantly dampen economic performance if they are proximate to major population and economic centers.

>>>>

What happens when cities themselves are threatened? In this paper, we bridge the economic development literature on the importance of cities with extant literature on the effect of armed conflict to provide a novel explanation for the paradox of high macroeconomic growth in conflict ridden countries. Specifically, we argue that to explain variation in economic performance between countries in the midst of civil war, it is necessary to account for the spatial location of the conflict relative to major urban centers. Much of the extant literature has focused on examining the effect of conflict on economic growth at the national level. The implicit assumption of these approaches is that all civil wars are alike, whether they occur right outside the capital or on the fringes of the country. We argue that the location of conflicts relative to major cities and capitals will determine the effect of civil conflict on economic performance, and that conflicts only significantly dampen economic performance if they are proximate to urban centers.
